Slavnov products and KP tau functions

### 2025-04-21 draft.tex
conclude first draft of the paper... refine it

Let us define the components
\begin{equation}
  \begin{split}
    \widetilde{\mathcal{K}}_{kj} = &
    \oint_{\gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    + \delta_{nj} F_k(\lambda) \\
    & = \sum_{\ell = 1}^n  \widehat{\Omega}_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    + F_k(\lambda)\delta_{nj}  \\
  \end{split}
\end{equation}
where
\begin{equation}
  \widehat{\Omega}_{k\ell}
  = \frac{1}{z_\ell} \frac{\widetilde{\mathcal{Y}}_k(\lambda, z_\ell)}{z_\ell - v_k}
  = \frac{(\lambda z_\ell -1)}{\lambda z_\ell} {\Omega}_{k\ell} 
\; , \qquad 
 F_k(\lambda) = (-1)^n \frac{\widehat{\mathcal{Y}}_k(\lambda; 0)}{v_k }  \prod_{s=1}^n z_s^{-1} \; .
\end{equation}

@@

@@

Now, we define the matrix \(\widetilde{\mathcal{K}}(\lambda)\) by its components
\begin{equation}
    \widetilde{\mathcal{K}}_{jk}(\lambda)  = -  \frac{1}{\lambda} 
  \oint_{\gamma_k} \frac{dw}{2\pi i} e^{\xi(\bm{t}, w^{- 1})} \frac{w^{-j -1 }\mathcal{Y}_k(w)}{w - v_k} \; ,
\end{equation}
and we then have 
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \bm{\tau}(\bm{t}, \bm{v})
  \det \left( \mathbb{1} +  \mathcal{K}^{-1} \widetilde{\mathcal{K}}(\lambda) \right)\; .
\end{equation}

We finally conclude that the Bakher-Akhiezer function can be written as
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \det \left( \mathbb{1} +  \mathcal{K}^{-1} \widetilde{\mathcal{K}}(\lambda) \right)\; .
\end{equation}

@@

It seems that we can consider a basis of functions \(\mathcal{Y}_j(z)\) for the Baker-Akhiezer
functions. I think that this object might say something interesting. In particular, we can
explicitly build solutions of the KP equation. In fact, the functions \(\mathcal{Y}_j\)
above have poles at the Bethe roots, it is interesting. 
\begin{itemize}
\item[\faCheck] What is the pole and root structure of the tau and Baker-Akhiezer (BA) functions?
  In particular, what happens at the Bethe roots?
\item[\faCheck] Basis tau functions
\item[\faCheck] Write the cases n=2 and n=3
\item[\faCheck] What the general case for all \(z_k\) approaching to a single
  value \(w\). 
\item[\faCheck] Write these tau functions in terms of Miwa coordinates
\item[\faCheck] I need to calculate the Baker-Akhiezer function for this model using the Japanese formula
  in terms of tau functions.
\item Write some explicit expressions for \(N=2\) and \(N=3\). 
\item Better investigate these functions. See Fredholm determinant. 
  I think that this BA function will be similar to the soliton-like solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}. 
  My functions seem to have a pole of order \(N\) at infinity. And \(N\) zeros at the Bethe roots. 
\item It turns out that the Baker-Akhiezer function is the appropriate object to discuss
  the Bethe roots. 
  \item I think that these tau functions come from algebro geometric tau functions. 
  \item I can try to study the dynamics closer to the Bethe roots, I think that the
    \(\Omega\)'s will have some interesting expansions, some of them might vanish. 
  \item This might become a Riemann Hilbert problem
  \item Are these tau functions associated to the Calogero-Moser model??
  \item Elliptic case - see slavnov, zabrodin's paper
\end{itemize}

>>>>

Let us define the components
\begin{equation}
  \begin{split}
    \widetilde{\mathcal{K}}_{kj} = &
    \oint_{\tilde \gamma_k} \frac{dw}{2\pi i} \prod_{s=1}^n \frac{1}{w - z_s}
    \frac{w^{n-j-1}\widehat{\mathcal{Y}}_k(w; \lambda)}{(w - v_k)}
    + \delta_{nj} F_k(\lambda) \\
    & = \sum_{\ell = 1}^n  \widehat{\Omega}_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{s\neq \ell} (z_\ell - z_s)} \right)
    + F_k(\lambda)\delta_{nj}  \\
  \end{split}
\end{equation}
where \(\tilde{\gamma}_k\) is a contour that does not pick
the point \(w=0\); and
\begin{equation}
  \widehat{\Omega}_{k\ell}
  = \frac{1}{z_\ell} \frac{\widetilde{\mathcal{Y}}_k(\lambda, z_\ell)}{z_\ell - v_k}
  = \frac{(\lambda z_\ell -1)}{\lambda z_\ell} {\Omega}_{k\ell} 
\; , \qquad 
 F_k(\lambda) = (-1)^n \frac{\widehat{\mathcal{Y}}_k(\lambda; 0)}{v_k }  \prod_{s=1}^n z_s^{-1} \; .
\end{equation}

@@

%%% Maybe I should calculate some examples
%\subsection{Examples} Let us now consider the cases \(n=2, 3\).
%We have calculated the tau functions for these objets. The
%elementary symmetric polynomials are also known. Therefore,
%the only detail we need to complete now is the calculation of
%the matrix \(\widehat{\mathcal{K}}\). 
%
%\subsubsection{n=2} We
%we have \(\mathcal{Y}_j\) for \(j=1,2\) and two Bethe roots

@@

We now want to consider the above calculations in the Miwa
coordinates. This allows us to consider the infinite chain limit.

We define the matrix \(\check{\mathcal{K}}(\lambda)\) by its components
\begin{equation}
    \check{\mathcal{K}}_{jk}  = -  
    \oint_{\gamma_k} \frac{dw}{2\pi i}
    e^{\xi(\bm{t}, w^{- 1})} \frac{w^{-j -1 }\mathcal{Y}_k(w)}{w - v_k} \; .
\end{equation}

Therefore, we have that 
\begin{equation}
  \widetilde{\mathcal{K}}_{kj} = 
  \mathcal{K}_{kj}  + \frac{1}{\lambda}
  \check{\mathcal{K}}_{kj}
  \ \Leftrightarrow \
  \widetilde{\mathcal{K}} = 
  \mathcal{K}  + \frac{1}{\lambda}
  \check{\mathcal{K}}\; , 
\end{equation}
and since \(\mathcal{K}\) is invertible, we get
\begin{equation}
  \widetilde{\mathcal{K}} = 
  \mathcal{K}\left( \mathbb{1}  +
  \frac{1}{\lambda}\mathcal{K}^{-1}\check{\mathcal{K}} \right)\; .
\end{equation}

Putting all these facts together, we find that the shifted
tau-functions are
\begin{equation}
  \bm{\tau}(\bm{t} - [\lambda^{-1}], \bm{v}) =
  \bm{\tau}(\bm{t}, \bm{v})
  \det \left( \mathbb{1} + \frac{1}{\lambda}
  \mathcal{K}^{-1} \check{\mathcal{K}} \right)\; .
\end{equation}

We finally conclude that the Bakher-Akhiezer function can be written as
\begin{equation}
  \psi(\bm{t},\bm{v}; \lambda) = e^{\xi(\bm{t}, \lambda)}
  \det \left( \mathbb{1} +
    \frac{1}{\lambda} \mathcal{T}
  \right)\; .
\end{equation}
where \(\mathcal{T} = \mathcal{K}^{-1} \check{\mathcal{K}}\).

@@

It seems that we can consider a basis of functions \(\mathcal{Y}_j(z)\) for the Baker-Akhiezer
functions. I think that this object might say something interesting. In particular, we can
explicitly build solutions of the KP equation. In fact, the functions \(\mathcal{Y}_j\)
above have poles at the Bethe roots, it is interesting. 
\begin{itemize}
\item[\faCheck] What is the pole and root structure of the tau and Baker-Akhiezer (BA) functions?
  In particular, what happens at the Bethe roots?
\item[\faCheck] Basis tau functions
\item[\faCheck] Write the cases n=2 and n=3
\item[\faCheck] What the general case for all \(z_k\) approaching to a single
  value \(w\). 
\item[\faCheck] Write these tau functions in terms of Miwa coordinates
\item[\faCheck] I need to calculate the Baker-Akhiezer function for this model using the Japanese formula
  in terms of tau functions.
\item[\faCheck]  Write some explicit expressions for \(N=2\) and \(N=3\). 
\item See the Limit \(L \to \infty\) and \(N \to \infty\) but sufficiently slower than \(L\). 
  I think that the matrix \(\mathcal{T}\) is a trace class operator and we write the BA function
  as a Fredholm determinant. 
\item Better investigate these functions in Miwa coordinates. See Fredholm determinant. 
  I think that this BA function will be similar to the soliton-like solution discussed in section 3.6.2 (page 41) of~\cite{Zabrodin2018}. 
  My functions seem to have a pole of order \(N\) at infinity. And \(N\) zeros at the Bethe roots. 
\item It turns out that the Baker-Akhiezer function is the appropriate object to discuss
  the Bethe roots. 
  \item I think that these tau functions come from algebro geometric tau functions. 
  \item I can try to study the dynamics closer to the Bethe roots, I think that the
    \(\Omega\)'s will have some interesting expansions, some of them might vanish. 
  \item This might become a Riemann Hilbert problem
  \item Are these tau functions associated to the Calogero-Moser model??
  \item Elliptic case - see slavnov, zabrodin's paper
\end{itemize}
